---
title: "GEO885 - MNF"
author: "Gregory Biland"
date: "3/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setenv(LANG = "en")

library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggmap)
library(flightplot)
library(sf)
library(shiny)
library(geosphere)
library(readxl)
library(slider)
library(scales)
library(wesanderson)
library(here)
library(maps)


# sf::sf_use_s2(FALSE)

dataFolder   <- here::here()   
RFolder      <- here::here()         
figureFolder <- here::here("figs") 

```

```{r Colors}
red <- wes_palette("Darjeeling1")[1]
green <- wes_palette("Darjeeling1")[2]
orange <- wes_palette("Darjeeling1")[3]
blue <- wes_palette("Darjeeling1")[5]
dark_orange <- wes_palette("Darjeeling1")[4]
black <- wes_palette("Darjeeling2")[5]
```


```{r Basic data preprocessing}
airports <- read.csv(file.path(dataFolder, "airports.csv"))

amm_incomplete <- read_excel(file.path(dataFolder, "MNF_amm.xlsx")) #Datensatz mit allen Flügen, vollständig und unvollständig

world <- st_read(file.path(dataFolder, "TM_WORLD_BORDERS_SIMPL-0 Kopie"))

```


### Data Transformation for emissions-API and flight data (pre processing)

```{r}
airports <- airports %>% dplyr::select(-c(continent, home_link, wikipedia_link, elevation_ft, municipality, scheduled_service, type, keywords, name, ident, iso_region, gps_code,local_code))
airports <- subset(airports, iata_code != "")

airports <- airports %>% st_as_sf(coords = c("longitude_deg", "latitude_deg"), crs = 4326, remove = FALSE) %>% rename(coordinates = geometry)
```

```{r Subset}
amm <- read.csv(file.path(dataFolder, "amm_complete.csv"))

# Alle Flugdaten mit "MISSING" values
amm <- subset(amm, select = -c(IATA_CODE_DEP))
amm <- subset(amm, select = -c(IATA_CODE_ARR))

amm$EMISSIONS_KGCO2EQ <- as.numeric(amm$EMISSIONS_KGCO2EQ)
amm <- amm %>% drop_na(EMISSIONS_KGCO2EQ)

paste("Completeness of dataset", (nrow(amm)/nrow(amm_incomplete))*100,"%")
```

### Link locations with coordinates
``` {r} 
amm_arr <- left_join(amm,airports, by = c("ARRIVAL_AIRPORT" = "iata_code"))
amm_arr <- amm_arr %>% rename(arr_coordinates = coordinates,
                      arr_country = iso_country)

amm <- left_join(amm,airports, by = c("DEPARTURE_AIRPORT" = "iata_code"))
amm <- amm %>% rename(dep_coordinates = coordinates,
                      dep_country = iso_country)
amm <- left_join(amm,airports, by = c("ARRIVAL_AIRPORT" = "iata_code"))
amm <- amm %>% rename(arr_coordinates = coordinates,
                      arr_country = iso_country)

drop <- c("id.x", "id.y")
amm = amm[,!(names(amm) %in% drop)]
```

```{r Distances}
amm <- amm %>% rename(dep_lon = longitude_deg.x,
                                        dep_lat = latitude_deg.x,
                                        arr_lon = longitude_deg.y,
                                        arr_lat = latitude_deg.y)

amm <- amm %>% rowwise() %>% 
    mutate(distance_km = (geosphere::distHaversine(c(dep_lon, dep_lat),
                                  c(arr_lon, arr_lat)))/1000)

emissions_per_class <- amm %>%
  group_by(year, cabin_class) %>%
  summarise(total_emissions = sum(EMISSIONS_KGCO2EQ))

aggregated_distance <- amm %>% 
  group_by(year) %>% 
  summarise(total_distance = sum(distance_km))
max_distance <- max(aggregated_distance$total_distance) 
```

```{r}
arrdep <- ggplot() +
  geom_sf(data=world)+
  geom_sf(data = amm, aes(geometry = dep_coordinates, col = "dep_coordinates"), size = .5, shape = 23) +
  geom_sf(data = amm, aes(geometry = arr_coordinates, col = "arr_coordinates"), size = .5, shape = 23) +
  ggtitle("Departure and arrival locations for all GIUZ flights separated by symbols")+
  xlab("Longitude") + ylab("Latitude")+
  scale_color_manual(labels = c("Arrival", "Departure"),
                     values = c("red", "blue"))+
  labs(colour = "Flight path")+
  theme(legend.position = "none")

arrdep
ggsave(file="flight_locations.png", arrdep,
       width = 10,
       height= 7)
```


```{r}
#https://www.gis-blog.com/flight-connection-map-with-r/

ZRH <- filter(amm, DEPARTURE_AIRPORT!="ZRH") 
#create basemap
map("world", fill=T, col="grey8", bg="grey15")
#overlay airports
points(amm$dep_lon, amm$dep_lat, pch=3, cex=0.1, col="chocolate1")+
points(amm$arr_lon, amm$arr_lat, pch=3, cex=0.1, col="chocolate1")

for (i in (1:dim(amm)[1])) { 
inter <- gcIntermediate(c(ZRH$dep_lon[1], ZRH$dep_lat[1]), c(amm$arr_lon[i], amm$arr_lat[i]), n=200)
lines(inter, lwd=0.1, col="turquoise2")    
}


```

```{r}
amm_arr <- st_as_sf(amm_arr)
# sf::sf_use_s2(FALSE)
flightmap <- st_join(world, amm_arr, join = st_contains_properly, left=TRUE)
flightmap_sort <- flightmap %>% group_by(NAME) %>% count() %>% filter(flightmap_sort$NAME != "Switzerland")

write.table(flightmap_sort,file="countryflights_counted.csv")
```

Subset all the flights that are entirely in Europa using the world- df and the amm- dataframe and crosslinking it bei ISO2 codes. 

```{r}
#ISO2 form european countries
EUISO2 <- read.csv(file.path(dataFolder, "EuroAlpha2-Codes.csv"), sep = ";") 
EUISO2$X <- NULL
EUISO2$ï..Name<-gsub("Â","",as.character(EUISO2$ï..Name)) 
names(EUISO2)[1] <- "Name"

#subset of european countries from world df
europe<- left_join(EUISO2, world)
europe$Name <- NULL
europe_ohne_RU <-subset(europe, ISO2!= "RU")
#just european flights arrival and departure 
europe_arr <- amm[which(amm$arr_country %in% unique(EUISO2$ISO2)),] 
europe_arr <- europe_arr %>% filter(arr_country != "RU")
europe_dep <- amm[which(amm$dep_country %in% unique(EUISO2$ISO2)),]
europe_dep <- europe_dep %>% filter(dep_country != "RU")

```

Create subset from world-df using the EUISO2-df and select all only European flights.
```{r}
arrdep_europe <-ggplot() +
  geom_sf(data=europe_ohne_RU, aes(geometry = geometry))+
  geom_sf(data = europe_dep, aes(geometry = dep_coordinates, col = "dep_coordinates"), size = 2, shape = 9) +
  geom_sf(data = europe_arr, aes(geometry = arr_coordinates, col = "arr_coordinates"), size = .75, shape = 19) +
  ggtitle("Departure and arrival locations for all european flights separated by symbols")+ xlab("Longitude") + ylab("Latitude")+
  scale_color_manual(labels = c("Arrival", "Departure"),
                     values = c("red", "blue"))+
  labs(colour = "Flight path") +
  theme(legend.position = "none")
arrdep_europe
```

```{r}
map("world", fill=T, col="grey8", bg="grey15")
#overlay airports
points(europe_dep$longitude_deg.x, europe_dep$latitude_deg.x, pch=3, cex=0.1, col="chocolate1")+
points(europe_arr$longitude_deg.y, europe_arr$latitude_deg.y, pch=3, cex=0.1, col="turquoise")
```

```{r}
cc_count_in <- amm_incomplete %>%
  group_by(cabin_class) %>%
  summarise(cnt = n()) %>%
  mutate(freq = round(cnt / sum(cnt), 2)) %>% 
  arrange(desc(freq))

cc_count_com <- amm %>%
  group_by(cabin_class) %>%
  summarise(cnt = n()) %>%
  mutate(freq = round(cnt / sum(cnt), 2)) %>% 
  arrange(desc(freq))
```


```{r compute baselines CO2}
reference_period <- c(2018:2020)
target_period <- c(2022, 2030)
full_period <- c(2018:2030)

aggregated_uzh <- amm %>% 
  group_by(year) %>% 
  summarise(total_emission = sum(EMISSIONS_KGCO2EQ, na.rm=TRUE)) %>%
  mutate(total_emission_3y_mean = slide_dbl(total_emission, mean, 
                                            .before = 1, .after = 1,
                                            .complete = T)) 
# UZH: only 2018 and 2019
uzh_baseline_co2 <- aggregated_uzh %>%
  filter(year %in% reference_period[]) %>%
  summarize(mean = mean(total_emission, na.rm=TRUE)) %>%
  pull(mean)

# Define target
uzh_target_co2 <- data.frame(x=target_period,
                             y=c(uzh_baseline_co2, 
                                 (uzh_baseline_co2/100)*0.47)) 

max_co2 <- max(aggregated_uzh$total_emission) 

```

```{r FLight Statistic}
# Total air travel at Giuz
total_co2_plot <- ggplot() + 
  geom_line(data=uzh_target_co2, 
            aes(x, y, color="target UZH"), linetype="dashed",  
            size=1) +
  geom_line(data=aggregated_uzh, 
            aes(x=year, y=total_emission, color="observed"),
            size=1) +
  ylab("kg CO2 eq.") +
  scale_x_continuous(breaks=full_period,
                     limits=c(min(full_period),
                              max(full_period))) +
  scale_y_continuous(label=comma,
                     limits=c(0,max_co2*1.02)) +
  ggtitle("UZH air travel (2018-2020): emissions")+
  scale_colour_manual("", 
                      breaks = c("observed","target UZH"),
                      values = c(green,orange))+
  theme_bw() +
  theme(text = element_text(size=16),
        axis.title.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = c(0.8, 0.25))

total_co2_plot
```


Difference in kgCO27km flying with economy and business class

```{r}
#Business
amm_business <- read.csv(file.path(dataFolder, "amm_business_complete.csv")) 
amm_business$EMISSIONS_KGCO2EQ <- as.numeric(amm_business$EMISSIONS_KGCO2EQ)
amm_business <- amm_business %>% drop_na(EMISSIONS_KGCO2EQ)

amm_business <- left_join(amm_business,airports, by = c("DEPARTURE_AIRPORT" = "iata_code"))
amm_business <- amm_business %>% rename(dep_coordinates = coordinates,
                      dep_country = iso_country)
amm_business <- left_join(amm_business,airports, by = c("ARRIVAL_AIRPORT" = "iata_code"))
amm_business <- amm_business %>% rename(arr_coordinates = coordinates,
                      arr_country = iso_country)
amm_business <- amm_business %>% rename(dep_lon = longitude_deg.x,
                                        dep_lat = latitude_deg.x,
                                        arr_lon = longitude_deg.y,
                                        arr_lat = latitude_deg.y)

amm_business <- amm_business %>% rowwise() %>% 
    mutate(distance_km = (geosphere::distHaversine(c(dep_lon, dep_lat),
                                  c(arr_lon, arr_lat)))/1000)
amm_business <- amm_business %>% rowwise() %>% mutate(kgCO2km = EMISSIONS_KGCO2EQ / distance_km)
amm_business = amm_business[,!(names(amm_business) %in% drop)]

#Economy
amm_economy <- read.csv(file.path(dataFolder, "amm_economy_complete.csv")) 
amm_economy$EMISSIONS_KGCO2EQ <- as.numeric(amm_economy$EMISSIONS_KGCO2EQ)
amm_economy <- amm_economy %>% drop_na(EMISSIONS_KGCO2EQ)

amm_economy <- left_join(amm_economy,airports, by = c("DEPARTURE_AIRPORT" = "iata_code"))
amm_economy <- amm_economy %>% rename(dep_coordinates = coordinates,
                      dep_country = iso_country)
amm_economy <- left_join(amm_economy,airports, by = c("ARRIVAL_AIRPORT" = "iata_code"))
amm_economy <- amm_economy %>% rename(arr_coordinates = coordinates,
                      arr_country = iso_country)

amm_economy <- amm_economy %>% rename(dep_lon = longitude_deg.x,
                                        dep_lat = latitude_deg.x,
                                        arr_lon = longitude_deg.y,
                                        arr_lat = latitude_deg.y)

amm_economy <- amm_economy %>% rowwise() %>% 
    mutate(distance_km = (geosphere::distHaversine(c(dep_lon, dep_lat),
                                  c(arr_lon, arr_lat)))/1000)

amm_economy <- amm_economy %>% rowwise() %>% mutate(kgCO2km = EMISSIONS_KGCO2EQ / distance_km)
amm_economy = amm_economy[,!(names(amm_economy) %in% drop)]
```

```{r}
eco2 <- ggplot(emissions_per_class, aes(fill=cabin_class, y=total_emissions, x=year)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_discrete(name = "Reason for travel")+
    ggtitle("Cabin class for each flight at MNF (2018-2020)") +
    xlab("") +
    ylab("emissions [kgCO2]") +
    scale_y_continuous(label=comma,
                       limits=c(0,max_distance*1.02)) +
    theme_bw() + 
    theme(legend.position = c(0.86, 0.7))
eco2

ggsave(file="flight_class.png", eco2,
       width = 10,
       height= 7)

eco2 <- ggplot(emissions_per_class, aes(fill=cabin_class, y=total_emissions, x=year)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_discrete(name = "Reason for travel")+
    ggtitle("Cabin class for each flight at MNF (2018-2020)") +
    xlab("") +
    ylab("emissions [kgCO2]") +
    scale_y_continuous(label=comma,
                       limits=c(0,max_distance*1.02)) +
    theme_bw() + 
    theme(legend.position = c(0.86, 0.7))
eco2

ggsave(file="flight_class.png", eco2,
       width = 10,
       height= 7)

eco2 <- ggplot(emissions_per_class, aes(fill=cabin_class, y=total_emissions, x=year)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_discrete(name = "Reason for travel")+
    ggtitle("Cabin class for each flight at MNF (2018-2020)") +
    xlab("") +
    ylab("emissions [kgCO2]") +
    scale_y_continuous(label=comma,
                       limits=c(0,max_distance*1.02)) +
    theme_bw() + 
    theme(legend.position = c(0.86, 0.7))
eco2

ggsave(file="flight_class.png", eco2,
       width = 10,
       height= 7)


```

