---
title: "GEO885 - MNF"
author: "Gregory Biland"
date: "3/7/2022"
output: html_document
---

```{r setup, include=FALSE}
Sys.setenv(LANG = "en")

library(knitr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggmap)
library(flightplot)
library(sf)
library(shiny)
library(geosphere)
library(readxl)
library(slider)
library(scales)
library(wesanderson)
library(here)
library(maps)
library(isoband)
library(zoo)
library(metR)
library(akima)
library(ggforce)
library(colorspace)

knitr::opts_chunk$set(
  echo = TRUE,
  message = F,
  fig.width = 7,
  fig.height = 6,
  pandoc.stack.size = "4g",
  fig.align = 'center',
  opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
)

# sf::sf_use_s2(FALSE)

dataFolder   <- here::here()   
RFolder      <- here::here()         
figureFolder <- here::here("figs") 

```

```{r Colors}
red <- wes_palette("Darjeeling1")[1]
green <- wes_palette("Darjeeling1")[2]
orange <- wes_palette("Darjeeling1")[3]
blue <- wes_palette("Darjeeling1")[5]
dark_orange <- wes_palette("Darjeeling1")[4]
black <- wes_palette("Darjeeling2")[5]
```


```{r Basic data preprocessing}
airports <- read.csv(file.path(dataFolder, "airports.csv"))

amm_incomplete <- read_excel(file.path(dataFolder, "MNF_amm.xlsx")) #Datensatz mit allen Flügen, vollständig und unvollständig

world <- st_read(file.path(dataFolder, "TM_WORLD_BORDERS_SIMPL-0 Kopie"))

employement_MNF <- data.frame(Jahr = c(2018,2019, 2020, 2021, 2022),
                              Personen = c(1538,1559,1656,1676,1651 ),
                              FTE = c(1119.0,1120.7,1170.7,1213.8,1190.3))
```


### Data Transformation for emissions-API and flight data (pre processing)

```{r}
airports <- airports %>% dplyr::select(-c(continent, home_link, wikipedia_link, elevation_ft, municipality, scheduled_service, type, keywords, name, ident, iso_region, gps_code,local_code))
airports <- subset(airports, iata_code != "")

airports <- airports %>% st_as_sf(coords = c("longitude_deg", "latitude_deg"), crs = 4326, remove = FALSE) %>% rename(coordinates = geometry)
```

```{r Subset}
amm <- read.csv(file.path(dataFolder, "amm_complete.csv"))

# Alle Flugdaten mit "MISSING" values
amm <- subset(amm, select = -c(IATA_CODE_DEP))
amm <- subset(amm, select = -c(IATA_CODE_ARR))

amm$EMISSIONS_KGCO2EQ <- as.numeric(amm$EMISSIONS_KGCO2EQ)
amm <- amm %>% drop_na(EMISSIONS_KGCO2EQ)

paste("Completeness of dataset", (nrow(amm)/nrow(amm_incomplete))*100,"%")
```

### Link locations with coordinates
``` {r} 
amm_arr <- left_join(amm,airports, by = c("ARRIVAL_AIRPORT" = "iata_code"))
amm_arr <- amm_arr %>% rename(arr_coordinates = coordinates,
                      arr_country = iso_country)

amm <- left_join(amm,airports, by = c("DEPARTURE_AIRPORT" = "iata_code"))
amm <- amm %>% rename(dep_coordinates = coordinates,
                      dep_country = iso_country)
amm <- left_join(amm,airports, by = c("ARRIVAL_AIRPORT" = "iata_code"))
amm <- amm %>% rename(arr_coordinates = coordinates,
                      arr_country = iso_country)

drop <- c("id.x", "id.y")
amm = amm[,!(names(amm) %in% drop)]
```

```{r Distances}
amm <- amm %>% rename(dep_lon = longitude_deg.x,
                                        dep_lat = latitude_deg.x,
                                        arr_lon = longitude_deg.y,
                                        arr_lat = latitude_deg.y)

amm <- amm %>% rowwise() %>% 
    mutate(distance_km = (geosphere::distHaversine(c(dep_lon, dep_lat),
                                  c(arr_lon, arr_lat)))/1000)

emissions_per_class <- amm %>%
  group_by(year, cabin_class) %>%
  summarise(total_emissions = sum(EMISSIONS_KGCO2EQ))

aggregated_distance <- amm %>% 
  group_by(year) %>% 
  summarise(total_distance = sum(distance_km))
max_distance <- max(aggregated_distance$total_distance) 
```

```{r}
arrdep <- ggplot() +
  geom_sf(data=world)+
  geom_sf(data = amm, aes(geometry = dep_coordinates, col = "dep_coordinates"), size = .5, shape = 23) +
  geom_sf(data = amm, aes(geometry = arr_coordinates, col = "arr_coordinates"), size = .5, shape = 23) +
  ggtitle("Departure and arrival locations for all GIUZ flights separated by symbols")+
  xlab("Longitude") + ylab("Latitude")+
  scale_color_manual(labels = c("Arrival", "Departure"),
                     values = c("red", "blue"))+
  labs(colour = "Flight path")+
  theme(legend.position = "none")

arrdep
ggsave(file="Figures/flight_locations.png", arrdep,
       width = 10,
       height= 7)
```


```{r}
#https://www.gis-blog.com/flight-connection-map-with-r/

ZRH <- filter(amm, DEPARTURE_AIRPORT!="ZRH") 
#create basemap
map("world", fill=T, col="grey8", bg="grey15")
#overlay airports
points(amm$dep_lon, amm$dep_lat, pch=3, cex=0.1, col="chocolate1")+
points(amm$arr_lon, amm$arr_lat, pch=3, cex=0.1, col="chocolate1")

for (i in (1:dim(amm)[1])){
inter <- gcIntermediate(c(amm$dep_lon[i], amm$dep_lat[i]), c(amm$arr_lon[i], amm$arr_lat[i]), breakAtDateLine = FALSE)
lines(inter, lwd=0.1, col="turquoise2") 
}

```

```{r include=FALSE}
countryflighty_counted <- read.csv(file.path(dataFolder, "countryflights_counted.csv"), sep = " ") #Without Switzerland
```

Subset all the flights that are entirely in Europa using the world- df and the amm- dataframe and crosslinking it bei ISO2 codes. 

```{r}
#ISO2 form european countries
EUISO2 <- read.csv(file.path(dataFolder, "EuroAlpha2-Codes.csv"), sep = ";") 
names(EUISO2)[1] <- "Name"
EUISO2$Name <-gsub("Â","",as.character(EUISO2$Name)) 


#subset of european countries from world df
europe<- left_join(EUISO2, world)
europe$Name <- NULL
europe_ohne_RU <-subset(europe, ISO2!= "RU")
#just european flights arrival and departure 
europe_arr <- amm[which(amm$arr_country %in% unique(EUISO2$ISO2)),] 
europe_arr <- europe_arr %>% filter(arr_country != "RU")
europe_dep <- amm[which(amm$dep_country %in% unique(EUISO2$ISO2)),]
europe_dep <- europe_dep %>% filter(dep_country != "RU")

europe_dep <- europe_dep %>% filter(DEPARTURE_AIRPORT != ARRIVAL_AIRPORT) %>% drop_na(distance_km)
europe_arr <- europe_arr %>% filter(DEPARTURE_AIRPORT != ARRIVAL_AIRPORT) %>% drop_na(distance_km)

dep_ZH <- europe_arr %>% filter(DEPARTURE_AIRPORT == "ZRH") #All flights origination in ZRH and landing in Europe

```

Create subset from world-df using the EUISO2-df and select all only European flights.
```{r}
arrdep_europe <-ggplot() +
  geom_sf(data=europe_ohne_RU, aes(geometry = geometry))+
  geom_sf(data = europe_dep, aes(geometry = dep_coordinates, col = "dep_coordinates"), size = 2, shape = 9) +
  geom_sf(data = europe_arr, aes(geometry = arr_coordinates, col = "arr_coordinates"), size = .75, shape = 19) +
  ggtitle("Departure and arrival locations for all european flights separated by symbols")+ xlab("Longitude") + ylab("Latitude")+
  scale_color_manual(labels = c("Arrival", "Departure"),
                     values = c("red", "blue"))+
  labs(colour = "Flight path") +
  theme(legend.position = "none")

arrdep_europe

ggsave(file="Figures/europe_flights.png", arrdep_europe,
       width = 10,
       height= 7)
```

```{r Modell 1}
interpolated <- interp(dep_ZH$arr_lon, dep_ZH$arr_lat, dep_ZH$EMISSIONS_KGCO2EQ, 
                       duplicate = "mean",
                       nx= 500,
                       ny = 500, jitter.random = TRUE)

#convert this to a long form dataframe
interp_df <- expand_grid(i = seq_along(interpolated$x), j = seq_along(interpolated$y)) %>% 
  mutate(lon = interpolated$x[i],
         lat = interpolated$y[j],
         emissions = map2_dbl(i, j, ~interpolated$z[.x,.y])) %>% 
  select(-i, -j) %>% data.frame()

#then you can use this in your plot
dep_ZH_plot_iso <- ggplot()+
  geom_sf(data=europe_ohne_RU, aes(geometry = geometry))+
  geom_contour_filled(data = interp_df, aes(x = lon, y = lat, z = emissions),show.legend = TRUE, alpha = .5)+
  scale_fill_viridis_d(option = "viridis",direction = -1)+
  geom_point(data = dep_ZH, aes(x = arr_lon, y = arr_lat), size = 1) +
  geom_point(data = dep_ZH, aes(x = dep_lon, y = dep_lat), col = "red", pch = 13, size = 2) +
  xlab("Longitude") + ylab("Latitude")+
  guides(fill=guide_legend(title="Emission zones [kgCO2]"))
  
dep_ZH_plot_iso

ggsave(file="Figures/dep_ZH_europeflights_iso.png", dep_ZH_plot_iso,
       width = 10,
       height= 7)
```

```{r Modell 2}

countryflighty_counted_EU <- left_join(europe_ohne_RU, countryflighty_counted) %>% drop_na()

dep_ZH_plot_distance <- ggplot()+
  geom_sf(data=countryflighty_counted_EU, aes(geometry = geometry, fill = n), alpha = .5)+
  scale_fill_viridis_c(option = "bilbao",direction = -1, labs( colour=  "Count per country"))+
  geom_point(data = dep_ZH, aes(x = dep_lon, y = dep_lat), col = "red", pch = 13, size = 3) +
  geom_point(data = dep_ZH, aes(x = arr_lon, y = arr_lat), size = 2, col = "black") +
  geom_circle(aes(y0 = 47.45806, x0 =8.548056, r = 11), geom = "circle", position = "identity", col = "red")+
  xlab("Longitude") + ylab("Latitude")

dep_ZH_plot_distance

ggsave(file="Figures/dep_ZH_europeflights_distance.png", dep_ZH_plot_distance,
       width = 10,
       height= 7)

```

```{r}
cc_count_in <- amm_incomplete %>%
  group_by(Serviceklasse) %>%
  summarise(cnt = n()) %>%
  mutate(freq = round(cnt / sum(cnt), 2)) %>% 
  arrange(desc(freq))

cc_count_com <- amm %>%
  group_by(cabin_class) %>%
  summarise(cnt = n()) %>%
  mutate(freq = round(cnt / sum(cnt), 2)) %>% 
  arrange(desc(freq))

amm <- amm %>% rowwise() %>% mutate(kgCO2km = EMISSIONS_KGCO2EQ / distance_km)
amm = amm[,!(names(amm) %in% drop)]
```


```{r compute baselines CO2}
amm_economy <- read.csv(file.path(dataFolder, "amm_economy_complete.csv"), sep = ",")
amm_business <- read.csv(file.path(dataFolder, "amm_business_complete.csv"), sep = ",")

reference_period <- c(2018:2020)
target_period <- c(2019, 2030)
full_period <- c(2018:2030)

aggregated_uzh_regular <- amm %>% 
  group_by(year) %>% 
  summarise(total_emission = sum(EMISSIONS_KGCO2EQ, na.rm=FALSE)) %>%
  mutate(total_emission_3y_mean = slide_dbl(total_emission, mean, 
                                            .before = 1, .after = 1,
                                            .complete = T)) 
#Determine difference in emissions between all economy and regular flight data
aggregated_uzh_economy <- amm_economy %>% 
  group_by(year) %>% 
  summarise(total_emission = sum(EMISSIONS_KGCO2EQ, na.rm=FALSE)) %>%
  mutate(total_emission_3y_mean = slide_dbl(total_emission, mean, 
                                            .before = 1, .after = 1,
                                            .complete = T)) 

#mean emissions for 2018 and 2019: 1384100
# UZH: only 2018 and 2019
uzh_baseline_co2_regular <- aggregated_uzh_regular %>%
  filter(year %in% reference_period[]) %>%
  summarize(mean = mean(total_emission, na.rm=TRUE)) %>%
  pull(mean)

uzh_baseline_co2_economy <- aggregated_uzh_economy %>%
  filter(year %in% reference_period[]) %>%
  summarize(mean = mean(total_emission, na.rm=TRUE)) %>%
  pull(mean)

# Define target
uzh_target_co2_regular <- data.frame(x=target_period,
                             y=c(uzh_baseline_co2_regular, 
                                 uzh_baseline_co2_regular*0.47))
uzh_target_co2_economy <- data.frame(x=c(2019,2019,2030),
                             y=c(uzh_baseline_co2_regular,uzh_baseline_co2_economy,uzh_baseline_co2_economy*0.49803))

max_co2 <- max(aggregated_uzh_regular$total_emission)

```

```{r}
#Tidy dataframes for kgCO2 per km analysis

amm <- amm %>% as.data.frame()
amm <- amm %>% filter_all(all_vars(!is.infinite(kgCO2km))) %>% drop_na(kgCO2km)

amm_business <- amm_business %>% filter(DEPARTURE_AIRPORT != ARRIVAL_AIRPORT)
amm_economy <- amm_economy %>% filter(DEPARTURE_AIRPORT != ARRIVAL_AIRPORT)
amm <- amm %>% filter(DEPARTURE_AIRPORT != ARRIVAL_AIRPORT)
```

Difference in kgCO27km flying with economy and business class

```{r}
#Plot actual emissions per cabin class for original dataset
eco2 <- ggplot() + 
    geom_col(data = emissions_per_class, aes(y=total_emissions, x=year,group = fct_rev(fct_inorder(cabin_class)), fill = cabin_class), position="stack", stat="identity", width = .5)+
    scale_fill_viridis_d(direction = -1, labs( colour=  "Cabin class"))+
    ggtitle("Cabin class for each flight at MNF (2018-2020)") +
    xlab("Year") +
    ylab("emissions [kgCO2]") +
    scale_y_continuous(label=comma,
                       limits=c(0,max_distance*1.02)) +
    theme_light() + 
    theme(legend.position = c(0.86, 0.7))
eco2

ggsave(file="Figures/flight_class.png", eco2,
       width = 10,
       height= 7)
```

```{r}
#Calculate emissions per cabinclass for altered data sets and merge with original to full list df
emissions_per_class_business <- amm_business %>%
  group_by(year, cabin_class) %>%
  summarise(total_emissions = sum(EMISSIONS_KGCO2EQ))

emissions_per_class_business$cabin_class[emissions_per_class_business$cabin_class == "business"] <- "all_business"

emissions_per_class_economy <- amm_economy %>%
  group_by(year, cabin_class) %>%
  summarise(total_emissions = sum(EMISSIONS_KGCO2EQ))

emissions_per_class_economy$cabin_class[emissions_per_class_economy$cabin_class == "economy"] <- "Model: all Economy"

#Transform year to observed and predicted
emissions_per_class_economy$year[emissions_per_class_economy$year == "2018"] <- "2018_predicted"
emissions_per_class_economy$year[emissions_per_class_economy$year == "2019"] <- "2019_predicted"
emissions_per_class_economy$year[emissions_per_class_economy$year == "2020"] <- "2020_predicted"

emissions_per_class$year[emissions_per_class$year == "2018"] <- "2018_observed"
emissions_per_class$year[emissions_per_class$year == "2019"] <- "2019_observed"
emissions_per_class$year[emissions_per_class$year == "2020"] <- "2020_observed"

emissions_per_class$cabin_class[emissions_per_class$cabin_class == "economy"] <- "Economy observed"
emissions_per_class$cabin_class[emissions_per_class$cabin_class == "business"] <- "Business observed"
emissions_per_class$cabin_class[emissions_per_class$cabin_class == "premium_economy"] <- "Premium economy observed"
emissions_per_class$cabin_class[emissions_per_class$cabin_class == "first"] <- "First observed"

emissions_per_class <- emissions_per_class %>% mutate(id = seq_along(year))
```


```{r}
vec_re <- c(2,3,1,5,6,4,8,10,7,9)

emissions_per_class <- emissions_per_class %>%
  slice(match(vec_re, id))

#Plot all three different data frames into one to compare emissions
all_emissions_plot <- ggplot() + 
    geom_col(data = emissions_per_class, aes(y=total_emissions, x=year,group = fct_rev(fct_inorder(cabin_class)), fill = cabin_class), position="stack", stat="identity", width = .5)+
    geom_col(data = emissions_per_class_economy, aes(y=total_emissions, x=year, fill = cabin_class), stat="identity", width = .5)+
    scale_fill_viridis_d(direction = -1, labs( colour=  "Cabin class"))+
    ggtitle("Emissions for each cabin class at the MNF (2018-2020)") +
    xlab("year") +
    ylab("emissions [kgCO2]") +
    scale_y_continuous(label=comma,
                       limits=c(0,max_distance*1.02)) +
    theme_light() 

all_emissions_plot
ggsave(file="Figures/emissions_comparison.png", all_emissions_plot,
       width = 10,
       height= 7)

label <- paste0("1000km threshold")

emission_distance_plot <- ggplot()+
  geom_point(data = amm, aes(y= kgCO2km, x = distance_km, col="blue"), size = .7, pch = 14)+
  geom_point(data = amm_economy, aes(y= kgCO2km, x = distance_km, col="green"), size = .4, pch = 23)+
  geom_vline(xintercept = 1000, linetype="dotted", size=.5)+
  scale_color_manual(labels = c("Observed", "Model: all Economy"),
                     values = c("440154FF", "5DC863FF"))+
  annotate("text", x = 2700, y = 1, label=label, color=black, size=3) + 
  ylab("kgCO2 emitted per flight km")+
  xlab("Flight distance [km]")+
  labs(colour = "Cabin class")+
  theme_light()

emission_distance_plot

ggsave(file="Figures/emissions_distance.png", emission_distance_plot,
       width = 10,
       height= 7)

```

# Reduction effect of all economy class flights per year to reach goal
```{r}
#mean all economy 1130033 kgCO2 ,6.25% reduction  per year

amm_shortdist <- amm_economy %>% filter(distance_km <= 1000)

mean(amm_shortdist$EMISSIONS_KGCO2EQ)

567787.5/226.7974
#2503.501
2503.501/nrow(amm_shortdist)
#0.9625148
#Reduce short dist flighty by 96% untill the year 2030

```

#Reduction goal untill 2030 562245.5kgCO2

```{r FLight Statistic}
label <- paste0("MNF reduction target")

# Total air travel at MNF
total_co2_plot <- ggplot() + 
  geom_line(data=uzh_target_co2_regular, 
            aes(x, y, color="target UZH"), linetype="dotted",  
            size=.75) +
  geom_line(data=uzh_target_co2_economy, 
            aes(x, y, color="target UZH all economy"), linetype="dotted",  
            size=.75) +
  # geom_line(data=aggregated_uzh_economy, 
  #           aes(x=year, y=total_emission, color="optimal"),
  #           size=.75) +
  # geom_line(data=aggregated_uzh_regular, 
  #           aes(x=year, y=total_emission, color="observed"),
  #           size=.75) +
  # geom_point(data=aggregated_uzh_economy, 
  #             aes(x=year, y=total_emission),
  #             color=orange, size=1.5) +
  # geom_point(data=aggregated_uzh_regular, 
  #             aes(x=year, y=total_emission),
  #             color=green, size=1.5) +
  geom_hline(yintercept = 562793.7, linetype="dotted", size=.5)+
  annotate("text", x = 2018.75, y = 590000, label=label, color=green, size=3) + 
  ylab("kg CO2 eq.") +
  scale_x_continuous(breaks=full_period,
                     limits=c(min(full_period),
                              max(full_period))) +
  scale_y_continuous(label=comma,
                     limits=c(0,max_co2*1.02)) +
  ggtitle("UZH air travel (2018-2020): emissions")+
  scale_colour_manual("",
                      breaks = c("target UZH", "target UZH all economy"),
                      values = c(red,blue))+
  theme_light() +
  theme(text = element_text(size=16),
        axis.title.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "right")

total_co2_plot

ggsave(file="Figures/target_goal_plot.png", total_co2_plot,
       width = 10,
       height= 7)
```

```{r}
#Statistics

paste("The mean emissions in kgCO2 per km flying for all business flights is", mean(amm_business$kgCO2km))
paste("The mean emissions in kgCO2 per km flying for all economy flights is", mean(amm_economy$kgCO2km))
paste("The mean emissions in kgCO2 per km flying for regular flights is", mean(amm$kgCO2km))

summary(amm)

```

